---
title: "CM515 Assignment 4"
author: "Melody Hayman"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

# Instructions

The aim of this assignment is to provide you with an opportunity to sharpen your skills in using ggplot(). While some of the plots you create may resemble those we covered in class, it is essential that they are original.

Rules for the Assignment:

1.All data sets except iris are fair game. That includes past datasets used in the class, sets built into R, your own data, and even online data.

2.  All graphs must include axis labels, plot title, a theme of your choice, and a brief description/interpretation of the plot (2-3 sentences).

It's worth noting that ggplot() is a widely-used tool, and there are numerous online resources available for you to explore. We strongly recommend that you take advantage of these resources to deepen your understanding of ggplot().

Knit the document into a PDF and submit it to canvas by 02/27/2024 at 11:59 pm.

# Load Packages and Data

------------------------------------------------------------------------

```{r}
####
#00#
####
##I just want to let you know I know this will not get full points. I was sick for this week and tried my best to figure all this out myself with the help of your Rmarkdowns, powerpoint, chatgpt, and google. Most of these I mostly understand now but some are still very confusing to me. I did the best I could while still turning it in on time. 


library(knitr)
library(tidyverse)


data("ToothGrowth")
ToothGrowth

#package in R, ToothGrowth, length of odontoblasts (cells for tooth growth) in 60 guinea pigs, three dose levels of vitamin C (0.5, 1, 2 mg/day) either by orange juice (OJ) or ascorbic acid (VC), C. I. Bliss (1952). The Statistics of Bioassay. Academic Press.






```

------------------------------------------------------------------------

# 1. Make a Scatter Plot With Customized Point Size and Transparency (3pts)

------------------------------------------------------------------------

Description and Interpretation: you can change the transparency (alpha) and size (size) both are set to represent length

```{r}
####
#01#
####


scatter_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, size = len, alpha = len)) +
  geom_point(color = "blue", shape = 16) +
  scale_size_continuous(range = c(3, 10)) +
  scale_alpha_continuous(range = c(0.5, 1)) +
  labs(title = "Tooth Growth Scatter Plot",
       x = "Dose",
       y = "Length") +
  theme_minimal()


print(scatter_plot)

```

------------------------------------------------------------------------

# 2. Fit a Line Through Your Scatter Plot From 1. (3pts)

------------------------------------------------------------------------

Description and Interpretation: geom_smooth adds a line

```{r}
####
#02#
####
scatter_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, size = len, alpha = len)) +
  geom_point(color = "blue", shape = 16) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +  
  scale_size_continuous(range = c(3, 10)) +
  scale_alpha_continuous(range = c(0.5, 1)) +
  labs(title = "Tooth Growth Scatter Plot with Fitted Line",
       x = "Dose",
       y = "Length") +
  theme_minimal()


print(scatter_plot)

```

# 3. Set the transparency, shape, and color of your graph from 1 based on expressions. Scale the color. (3pts)

------------------------------------------------------------------------

Description and Interpretation: This one really confused me and I could not get it to work

```{r}
####
#03#
####

transparency_expr <- expression(len / max(len))
shape_expr <- expression(ifelse(dose > 1, 17, 16))  
color_expr <- expression(dose * 10)  


scatter_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, size = len)) +
  geom_point(aes(alpha = transparency_expr, shape = shape_expr, color = color_expr)) +
  scale_size_continuous(range = c(3, 10)) +
  scale_alpha_continuous(range = c(0.5, 1)) +
  scale_color_gradient(low = "blue", high = "red") +  
  labs(title = "Tooth Growth Scatter Plot with Fitted Line",
       x = "Dose",
       y = "Length") +
  theme_minimal()


print(scatter_plot)

```

------------------------------------------------------------------------

# 4. Make a Boxplot With Customized Boxplot Width (3pts)

------------------------------------------------------------------------

Description and Interpretation: can adjust width with geom_boxplot(width = )

```{r}
####
#04#
####
boxplot <- ggplot(ToothGrowth, aes(x = as.factor(dose), y = len, fill = as.factor(dose))) +
  geom_boxplot(width = 0.5) +  
  labs(title = "Tooth Growth Data Boxplot",
       x = "Dose",
       y = "Length") +
  theme_minimal()


print(boxplot)

```

------------------------------------------------------------------------

# 5. Overlay the Individual Points Over Your Box Plot From 3. and Adjust the Point Size and Transparency as Needed (3pts)

------------------------------------------------------------------------

Description and Interpretation: not sure why this isnt working

```{r}
####
#05#
####
scatter_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, size = len, alpha = len)) +
  geom_point(color = "blue", shape = 16) +
  scale_size_continuous(range = c(3, 10)) +
  scale_alpha_continuous(range = c(0.5, 1)) +
  labs(title = "Tooth Growth Scatter Plot",
       x = "Dose",
       y = "Length") +
  theme_minimal()


boxplot <- ggplot(ToothGrowth, aes(x = as.factor(dose), y = len, fill = as.factor(dose))) +
  geom_boxplot(width = 0.5, alpha = 0.5) +
  labs(title = "Tooth Growth Data Boxplot",
       x = "Dose",
       y = "Length") +
  theme_minimal()


combined_plot <- scatter_plot + boxplot


print(combined_plot)

```

# 6. Make a plot and modify the scale in a beneficial way using scale_x_continuous(), scale_y_continuous(), and coordinate_cartesian(). Explain how your modifications to the scales improved the plot. (4pts)

------------------------------------------------------------------------

Description and Interpretation:This also confused me, Im not sure why this is not working. modifications to the scale can improve the focus on the data, customize the tick marks and adjust the axis limtis. These call all improve the readability of the graph.

```{r}
####
#06#
####
transparency_expr <- expression(len / max(len))


scatter_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, size = len, alpha = transparency_expr)) +
  geom_point(color = "blue", shape = 16) +
  scale_size_continuous(range = c(3, 10)) +
  scale_alpha_continuous(range = c(0.5, 1)) +
  labs(title = "Tooth Growth Scatter Plot",
       x = "Dose",
       y = "Length") +
  theme_minimal()


modified_scatter_plot <- scatter_plot +
  scale_x_continuous(limits = c(0, 2), breaks = seq(0, 2, 0.5)) +  
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, 10)) +  
  coord_cartesian(xlim = c(0, 2), ylim = c(0, 40))

print(modified_scatter_plot)
```

------------------------------------------------------------------------

# 7. Make a New (don't use any plots from 1-4) Plot with facet_wrap() (4pts)

------------------------------------------------------------------------

Description and Interpretation: Using facet_wrap we can see the realtionship between more then just two variables, so we can see if there are any patterns between length, dose, and supplemental type.

```{r}
#07#
####
facet_lineplot <- ggplot(ToothGrowth, aes(x = as.factor(dose), y = len, group = supp, color = supp)) +
  geom_line() +
  labs(title = "Tooth Growth Line Plot with Facet Wrap",
       x = "Dose",
       y = "Length") +
  facet_wrap(~supp, ncol = 2) +  
  theme_minimal()


print(facet_lineplot)

```

# 8. Using the plot from 5, scale the colors and reorder your facet to reverse order. (4pts)

------------------------------------------------------------------------

Description and Interpretation: that graph was not working so I used my basic scatterplot that did work

```{r}
#08#
####
scatter_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, size = len, alpha = len, color = supp)) +
  geom_point(shape = 16) +
  scale_size_continuous(range = c(3, 10)) +
  scale_alpha_continuous(range = c(0.5, 1)) +
  scale_color_manual(values = c("blue", "red")) + 
  labs(title = "Tooth Growth Scatter Plot",
       x = "Dose",
       y = "Length") +
  theme_minimal() +
  facet_wrap(~supp, scales = "free_x", ncol = 2)  


print(scatter_plot)

```

------------------------------------------------------------------------

# 9. Make a Plot Using a Geometry That Was Not Covered in Class (4pts)

------------------------------------------------------------------------

Description and Interpretation:

```{r}
####
#09#
####
violin_plot <- ggplot(ToothGrowth, aes(x = as.factor(dose), y = len, fill = supp)) +
  geom_violin() +
  labs(title = "Violin Plot of Tooth Growth",
       x = "Dose",
       y = "Length") +
  theme_minimal()
print(violin_plot)
```

# 10. Make a Plot Using a ggplot Extension That Was Not Covered in Class. Explain what the extension does. (4pts)

------------------------------------------------------------------------

Description and Interpretation:This extension lets you make a beeswarm plot. This is a type of scatter plot where categories are placed without overlapping on the x axis. This allows for another way to look at the values of each dose level compared to each other in respect to length.

```{r}
####
#10#
####

#install.packages("ggbeeswarm")
library(datasets)
library(ggplot2)
library(ggbeeswarm)


beeswarm_plot <- ggplot(ToothGrowth, aes(x = as.factor(dose), y = len, color = supp)) +
  geom_beeswarm() +
  labs(title = "Beeswarm Plot of Tooth Growth",
       x = "Dose",
       y = "Length") +
  theme_minimal()


print(beeswarm_plot)
```

------------------------------------------------------------------------

# Appendix

```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}

```
